---
title: "naiveBayes Case Study"
author: "Claus Lv"
date: "Thursday, September 18, 2014"
output: html_document
---

## Naïve Bayes


## Available packages on CRAN
`e1071` and `klaR`

```{r load library}
library(e1071)
```

## Iris Data Set
The Iris dataset contains 150 instances, corresponding to three equally-frequent species of iris plant (Iris setosa, Iris versicolour, and Iris virginica). An Iris versicolor is shown below, courtesy of Wikimedia Commons.

```{r Iris data set summary}
summary(iris)
pairs(iris[1:4], main = "Iris Data (red=setosa,green=versicolor,blue=virginica)",
      pch = 21,
      bg = c("red", "green3", "blue")[unclass(iris$Species)])
```

## Classify use naive Bayes
```{r analysis}
## column from 1 to 4 is the attributes, col 5 is the class
classifier <- naiveBayes(iris[ ,1:4], iris[,5])
## result
table(predict(classifier, iris[,-5]), iris[,5])
```

This simple case study shows that a Naïve Bayes classifier makes few mistakes in a dataset that, although simple, is not linearly separable, as shown in the scatterplots and by a look at the confusion matrix, where all misclassifications are between Iris Versicolor and Iris Virginica instances.

## Note
